---
title: "Assignment 0 - Getting started"
author: "ADITYA NARAYAN RAI + adityanarayan-rai"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---
  
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

<!-- Do not forget to input your Github username in the YAML configuration up there --> 

***

```{r, include = T}
# LOAD THE PACKAGES YOU ARE USING IN THIS CODE CHUNK 
library(tidyverse)
library(ggplot2)
library(dplyr)
```


### Task 1 - Dealing with a numeric vector


Consider the following numeric vector.

```{r}
x <- c(34, 56, 55, 87, NA, 4, 77, NA, 21, NA, 39)
```

a)  What's the mean of all values? And what's the median?

```{r}
mean(x, na.rm = T)
median(x, na.rm = T)
```

b) What's the variance and the standard deviation of all values?

```{r}
var(x, na.rm = T)
sd(x, na.rm = T)
```

c) Compute the sum of all values in the vector `x`.
```{r}
sum(x, na.rm = T)
```

d) Create a new vector, `x_std`, that standardizes `x`, i.e. transforms the values in a way that `mean(x_std) = 0` and `sd(x_std) = 1`.

```{r}
x_std <- (x - mean(x, na.rm = T))/sd(x, na.rm = T)
mean(x_std, na.rm = T)
sd(x_std, na.rm = T)
```

<br>

***

### Task 2 - Dealing with a character vector

Consider the following character vector.

```{r}
friends <- c("Homer", "Marge", "Bart", "Lisa", "Maggie")
```


a) Get the length of the vector.

```{r}
length(friends)
```

b) Extract the first two elements of the vector.

```{r}
friends[c(1,2)]
```

c) Extract the second and the last observation from the vector.

```{r}
friends[c(2,5)]
```

d) Sort the elements in reverse alphabetic order.

```{r}
sort(friends, decreasing = TRUE)
```


<br>

***

### Task 3 - A dataset on immigration trips to Ellis Island

a) Load the dataset `immigration-ellis.tsv` into R, which contains data on immigration to Ellis Island (1918-1922) by trip.

```{r}
immigration <- read.delim("immigration-ellis.tsv")
```

b) What is the total number of immigration cases in the dataset?

```{r}
nrow(immigration)
```

c) From where did the ship with the most immigrants arrive?

```{r}
most_immigrants_ship <-immigration[which.max(immigration$Count), ]
most_immigrants_ship
```

d) How many unique ships (according to the name) are in the dataset?

```{r}
unique_ships <- unique(immigration$Ship)
num_unique_ships <- length(unique_ships)
num_unique_ships
```

e) Find out the name of the package from the tidyverse that makes working with date and times data easier, and report it below!
lubridate

<!-- During this course, we will input our answers that include text inside the tag, as such:

<div class = "answer"> 
YOU PUT YOUR TEXT HERE
</div> 
 
Also, you will notice that the text you put inside here will not appear in your .html file.
That is because very similar to the # in our R scripts, this is a comment tag for html.
-->

<div class = "answer">

lubridate

</div>

